```{r}
#| warning: False
lapply(c("httr2","sf","tidyverse","ows4R","rpostgis","XML"), require, character.onl=TRUE)
```

## Import data
sudo apt install tesseract-ocr-eng
Error opening data file /usr/share/tesseract-ocr/4.00/tessdata/eng.traineddata
Please make sure the TESSDATA_PREFIX environment variable is set to your "tessdata" directory.
Failed loading language 'eng'
Tesseract couldn't load any languages!
Error: Unable to find training data for: eng. Please consult manual for: ?tesseract_download
### Insignia
#### Geojson
#### PostGIS
#### WFS
```{r}
roads<- read_sf("/media/ricardo/STORAGE_RicardoRS_A/MasterUNED/TFM/RT_MALAGA/RT_VIARIA/rt_tramo_vial.shp")
sf::st_write(roads, "roads_es.geojson")
wfs_nuts_es <- "https://www.ign.es/wfs-inspire/unidades-administrativas?request=GetCapabilities&service=WFS"

wfs_nuts <- "https://openiacs.eu/geoserver/ows?service=wfs&request=GetCapabilities#iacs:pg_nuts_00"
wfs_nuts_es_url <- url_parse(wfs_nuts)
wfs_nuts_eu_client <- ows4R::WFSClient$new(wfs_nuts,
                                    serviceVersion = "2.0.0")
wfs_nuts_eu_client$getFeatureTypes(pretty=TRUE)
wfs_nuts_eu_client$describeFeatureType(typeName="au:AdministrativeBoundary")
wfs_nuts_es_url$query <- list(service = "WFS",
                  version = "2.0.0",
                  request = "GetFeature",
                  typename = "iacs:pg_nuts_03",
                  outputFormat = "application/json")
drenthe_sf <- wfs_nuts_es_url |>
  url_build() |> 
  st_read()
agp <- drenthe_sf[str_detect(drenthe_sf$nuts_name,"Málaga"),]

st_inter
```

### MITECO
#### Geojson

```{r}
st_layers <- "/home/ricardo/Downloads/094-dera-11-patrimonio-gpkg-6f7g/11_Patrimonio.gpkg"
patrimonio<- read_sf("/home/ricardo/Downloads/094-dera-11-patrimonio-gpkg-6f7g/11_Patrimonio.gpkg",layer="T11_07_Enp_FiguraProteccion")

WFSClient$new(url=)
```

#### PostGIS
```{r}
conn <- RPostgreSQL::dbConnect("PostgreSQL", host="localhost",
                       dbname="rruiz_s_gis",
                       user="ricardo",
                       password="Rrs90horn190")
eennpp_andalucia_postgis <- read_sf(conn, layer="patrimonio_proteccion")
RPostgreSQL::dbGetQuery(conn, "SELECT * FROM patrimonio_proteccion * WHERE ID ='32'")
```
#### WFS


### NUTS - GiscoR

```{r}
wfs_nuts_es <- "https://www.ign.es/wfs-inspire/unidades-administrativas?request=GetCapabilities&service=WFS"

wfs_nuts <- "https://openiacs.eu/geoserver/ows?service=wfs&request=GetCapabilities#iacs:pg_nuts_00"
wfs_nuts_es_url <- url_parse(wfs_nuts)
wfs_nuts_eu_client <- ows4R::WFSClient$new(wfs_nuts,
                                    serviceVersion = "2.0.0")
wfs_nuts_eu_client$getFeatureTypes(pretty=TRUE)
wfs_nuts_eu_client$describeFeatureType(typeName="au:AdministrativeBoundary")
wfs_nuts_es_url$query <- list(service = "WFS",
                  version = "2.0.0",
                  request = "GetFeature",
                  typename = "iacs:pg_nuts_03",
                  outputFormat = "application/json")
drenthe_sf <- wfs_nuts_es_url |>
  url_build() |> 
  st_read()
Málaga <- drenthe_sf[str_detect(drenthe_sf$nuts_name,"Málaga"),]
```


```{r}
wfs_natura_andalucia <- "http://www.juntadeandalucia.es/medioambiente/mapwms/REDIAM_WFS_Patrimonio_Natural"
wfs_natura_client <- ows4R::WFSClient$new(wfs_natura_andalucia,
                                             serviceVersion = "2.0.0")
# all "layers"
wfs_natura_client$getFeatureTypes(pretty=TRUE)
# all fields from ms:EENNPP
wfs_natura_client$describeFeatureType(typeName = "ms:EENNPP") |>
  map_chr(function(x){x$getName()})
# possibilities to extract
wfs_natura_client$
  getCapabilities()$
  getOperationsMetadata()$
  getOperations() %>%
  map(function(x){x$getParameters()}) %>%
  pluck(3, "outputFormat")
## all EENNPP
url <- url_parse(wfs_natura_andalucia)
url$query <- list(service ="wfs",
                  request="GetFeature",
                  )


request <- url_build(url)
```

```{r}
gemeente_url <- url_parse("http://www.juntadeandalucia.es/medioambiente/mapwms/REDIAM_WFS_Patrimonio_Natural")
filter_drenthe <- "<PropertyName>NOMBRE</PropertyName><Literal>CÁRCAVAS DE MARCHAL</Literal>"
gemeente_url$query <- list(service = "WFS",
                  version = "2.0.0",
                  request = "GetFeature",
                  typename = "ms:EENNPP",
                  outputFormat = "application/json")

drenthe_sf <- gemeente_url |>
  url_build() |> 
  st_read()
```


```{r}
gemeente_url <- url_parse("https://service.pdok.nl/kadaster/bestuurlijkegebieden/wfs/v1_0")
filter_drenthe <- "<PropertyName>ligtInProvincieNaam</PropertyName><Literal>Drenthe</Literal>"
gemeente_url$query <- list(service = "WFS",
                  version = "2.0.0",
                  request = "GetFeature",
                  typename = "bestuurlijkegebieden:Gemeentegebied",
                  filter = paste0("<Filter><PropertyIsEqualTo>", filter_drenthe, "</PropertyIsEqualTo></Filter>"),
                  outputFormat = "application/json")

drenthe_sf <- gemeente_url |>
  url_build() |> 
  st_read()
```

## Ruta GPS
```{r}
gpx_parsed <- htmlTreeParse("~/quarto_github_rruiz_s/rruiz-s.github.io/data/ruta-18-rio-bermuza.gpx", useInternalNodes = TRUE)
coords <- xpathSApply(doc = gpx_parsed, path = "//trkpt", fun = xmlAttrs)
elevation <- xpathSApply(doc = gpx_parsed, path = "//trkpt/ele", fun = xmlValue)

df <- data.frame(
  lat = as.numeric(coords["lat", ]),
  lon = as.numeric(coords["lon", ]),
  elevation = as.numeric(elevation)
)

head(df, 10)
tail(df, 10)

plot(x = df$lon, y = df$lat, type = "l", col = "black", lwd = 3,
     xlab = "Longitude", ylab = "Latitude")

library(leaflet)

leaflet() %>%
  addTiles() %>%
  addPolylines(data = df, lat = ~lat, lng = ~lon, color = "#000000", opacity = 0.8, weight = 3)
```

## Transforming


# pdf

